\pkg{Rgbp}: An R Package for Gaussian, Poisson, and Binomial Hierarchical Modeling

### 2013-07-29 jss/article.tex
proofreading

where $m$ is the number of regression coefficients. As for $\beta$, it is a reasonable choice to assume a flat (non-informative) distribution because information about the second-level mean gets plentiful as the number of groups (\emph{k}) increases. The next flat prior distribution of the second-level variance component $A$ (or $1/r$) is chosen for its good repeated sampling properties for point and interval estimates of the true parameter $\{\mu_{j}\}$ over 2-level model \citep{2011}, and for posterior propriety under moderate conditions.

@@

\subsubsection{Simple Unbiased Estimates of Coverage Probabilities}
Let's define an indicator variable, $I^{(i)}_{j}$, which is 1 if the $j$-th group's interval estimate from the $i$-th pseudo-dataset includes $\mu^{(i)}_{j}$ and 0 otherwise. Then \code{coverage} estimates the coverage probability via the sample mean of these indicator variables for each group, $j$, averaging over all the possible $\mu^{(i)}_{j}$ and $y^{(i)}_{j}$ for given $A$ and $\beta$. This provides it a simple unbiased estimate of the $k$ coverage probabilities. For example, $\frac{1}{500}\sum_{i=1}^{500}I^{(i)}_{1}$ is the simple unbiased estimate of coverage probability for the first group ($j=1$), accounting for possible $\mu^{(i)}_{1}$ and $y^{(i)}_{1}$, $i=1, \ldots, 500,$ given a specific $(A, \beta)$. As the number of pseudo-datasets increases, this simple estimate converges to the true coverage probability of the first group given $A$ and $\beta$.

@@

For this purpose, data was gathered from 31 hospitals in NY composed of the number of deaths ($z$) within a month of CABG surgeries and total number of patients ($n$) receiving CABG surgeries in each hospital. And \pkg{Rgbp} will provide point and interval estimates of the true death ratio for each hospital so that one can choose the safest hospital. The following code is an example of input based on the last ten hospital data ($j=22, 23, \ldots, 31$).
\begin{CodeChunk}
\begin{CodeInput}
R> z <- c( 14,   9,  15,  13,  35,  26,  25,  20,   35,   27)
R> n <- c(593, 602, 629, 636, 729, 849, 914, 940, 1193, 1340)
\end{CodeInput}
\end{CodeChunk}

@@

Intuitively, the result of multilevel modeling makes more sense than that of naive sample mean. For example, suppose there are two hospitals, whose observed death ratios ($y_{j}$) are 0 and 0.001 and caseloads ($n_{j}$) are 1 and 1000 each. Do you believe that the former hospital, whose observed death rate is 0, is better than the latter and are you going to choose the former hospital? Borrowing information from state-level hierarchy seems reasonable for the former hospital because it is hard to judge its true death rate per exposure with just one caseload. Though somewhat extreme, this is what happened to the two left-most hospitals in the first plot and this is why hierarchical modeling is a reasonable choice for this dataset.

@@

\section[Discussion]{Discussion}
\pkg{Rgbp} is an R package for estimating and validating a two-level hierarchical model. The package aims to provide a procedure that is not only fast and easy to use but has good frequency properties and can be used in many scenarios. The package provides ``frequency method checking'' functionality to examine repeated sampling properties and test that the method is valid at a wide range of parameter values.

>>>>

where $m$ is the number of regression coefficients. As for $\beta$, it is a reasonable choice to assume a flat (non-informative) distribution because information about the second-level mean gets plentiful as the number of groups (\emph{k}) increases. The next flat prior distribution of the second-level variance component $A$ (or $1/r$) is chosen for its good repeated sampling properties for point and interval estimates of the true parameters $\{\mu_{j}\}$, $\{\lambda_{j}\}$, or $\{p_{j}\}$ over 2-level model \citep{2011}, and for posterior propriety under moderate conditions.

@@

\subsubsection{Simple Unbiased Estimates of Coverage Probabilities}
Let's define an indicator variable, $I^{(i)}_{j}$, which is 1 if the $j$-th group's interval estimate from the $i$-th pseudo-dataset includes $\mu^{(i)}_{j}$ and 0 otherwise. Then \code{coverage} estimates the coverage probability via the sample mean of these indicator variables for each group $j$, averaging over all the possible $\mu^{(i)}_{j}$ and $y^{(i)}_{j}$ for given $A$ and $\beta$. This provides a simple unbiased estimate of the $k$ coverage probabilities. For example, $\frac{1}{500}\sum_{i=1}^{500}I^{(i)}_{1}$ is the simple unbiased estimate of coverage probability for the first group ($j=1$), accounting for possible $\mu^{(i)}_{1}$ and $y^{(i)}_{1}$, $i=1, \ldots, 500,$ given a specific $(A, \beta)$. As the number of pseudo-datasets increases, this simple estimate converges to the true coverage probability of the first group given $A$ and $\beta$.

@@

For this purpose, data were gathered from 31 hospitals in NY composed of the number of deaths ($z$) within a month of CABG surgeries and total number of patients ($n$) receiving CABG surgeries in each hospital. Using these data, \pkg{Rgbp} provides point and interval estimates of the true death ratio for each hospital so that one can choose the safest hospital. The following code is an example of input based on the last ten hospital data ($j=22, 23, \ldots, 31$).
\begin{CodeChunk}
\begin{CodeInput}
R> z <- c( 14,   9,  15,  13,  35,  26,  25,  20,   35,   27)
R> n <- c(593, 602, 629, 636, 729, 849, 914, 940, 1193, 1340)
\end{CodeInput}
\end{CodeChunk}

@@

Intuitively, the result of multilevel modeling makes more sense than that of naive sample mean. For example, suppose there are two hospitals, whose observed death ratios ($y_{j}$) are 0 and 0.001 and caseloads ($n_{j}$) are 10 and 1000 each. Do you believe that the former hospital, whose observed death rate is 0, is better than the latter and are you going to choose the former hospital? Borrowing information from state-level hierarchy seems reasonable for the former hospital because it is hard to judge its true death rate per exposure with just ten caseloads. Though somewhat extreme, this is what happened to the two left-most hospitals in the first plot and this is why hierarchical modeling is a reasonable choice for this dataset.

@@

\section[Discussion]{Discussion}
\pkg{Rgbp} is an R package for estimating and validating a two-level hierarchical model. The package aims to provide a procedure that is not only fast and easy to use but has good frequency properties and can be used in many scenarios. The package provides ``frequency method checking'' functionality to examine repeated sampling properties and test that the method is valid at given hyper-parameter values.
